OpenMPIR

### 2017-09-08 paper.tex
More citations

When writing task-parallel programs today, one has a large selection of
potential programming models and implementations to consider. Unfortunately, 
despite having some significant overlaps in semantics, parallel programming models 
such as OpenMP\cite{}, Cilk\cite{},Kokkos\cite{}, HPX\cite{}, Charm++\cite{},
Qthreads\cite{}, pthreads\cite{}, MPI\cite{}, Chapel\cite{}, UPC\cite{},
among others have limited ability to interoperate either at the level of 
compile-time analysis or at run-time execution.
\wmnote{Not sure if relevant, but there was some experiment run a while back of someone composing a library that internally used OpenMP
with another framework that had terribly results compared with either used alone.} 
\gsnote{Hmm. A quick googling turned up a project out of berkeley combining OpenMP and TBB, 
but performance didn't seem too bad. I'll take another look tomorrow.}
In addition, many of these programming models are implemented as runtime libraries with only
primitive compiler support. As a result, most are not able to take advantage of the compiler's 
analysis and optimization capabilities.
While there has been some recent work on specializing compilers to reason
about parallel programming libraries~\cite{Moss_2016}, fragmentation among
models presents a significant challenge for compiler writers, which must choose
a single model, or multiple intermediate forms, to use for analysis and optimization. An ideal 
solution would be a form of common representation of parallel semantics and constructs.

@@

Arguably the most ubiquitous parallel programming model for on-node parallelism
in HPC, OpenMP has grown from being a way to implement parallel loops into a
large and complex specification for parallelism. This paper focuses on a 
relatively recent addition to the OpenMP specification: the \texttt{task}
and \texttt{taskwait} constructs. A detailed description of these constructs
is presented in Section~\ref{Sec:OpenMP} and a more complete description can 
be found in the OpenMP specification~\cite{spec}.

@@

In this section we further discuss a significant challenge in HPC:
fragmentation of parallel programming models. There are many ways to write
parallel programs~\cite{spec, qthreads, chapel, cilk, kokkos, legion} \pmnote{I
would add MPI and UPC/PGAS here too}. From
libraries, to language extensions, to standalone languages, to combinations of
the above, fragmentation of parallel programming models is a problem that is
only getting worse. This means that when writing tooling, optimizations, or
analyses, one has to choose which model to target as well as understand points
of contention and interoperability between them. A common difficulty among
all of these is that reasoning about parallel programs is hard and compilers
can potentially help reduce this burden. This is exactly analogous the problem 
LLVM solves: compiler optimizations are hard, so sharing them is valuable. Tapir 
is a demonstration that the the LLVM approach to compilation of serial code can 
be effectively extended to help reason about parallel code.

@@

Initially, the cross-vendor OpenMP~\cite{spec} shared memory programming model 
focused on the execution of data parallelism by a cooperating team of threads, 
e.g., dividing the iterations of a loop among the threads. Version 3.0 of the 
OpenMP API specification introduced support for lightweight asynchronous tasks, 
designated by the application developer and scheduled onto the team of threads 
by the OpenMP run time implementation.  The \texttt{task} construct applied to 
a structured block of code creates an explicit task, and the \texttt{taskwait} 
construct waits for completion of all tasks generated by the current task.

@@

>>>>

When writing task-parallel programs today, one has a large selection of
potential programming models and implementations to consider. Unfortunately, 
despite having some significant overlaps in semantics, parallel programming
models such as OpenMP~\cite{openmp}, Cilk~\cite{cilk}, Kokkos~\cite{kokkos},
HPX~\cite{hpx}, Charm++~\cite{charm}, Qthreads~\cite{qthreads},
pthreads\cite{pthreads}, MPI~\cite{mpi}, Chapel~\cite{chapel}, UPC~\cite{upc},
OpenCL~\cite{opencl}, OpenACC~\cite{openacc} among others have limited ability to
interoperate either at the level of compile-time analysis or at run-time
execution.  \wmnote{Not sure if relevant, but there was some experiment run a
while back of someone composing a library that internally used OpenMP with
another framework that had terribly results compared with either used alone.}
\gsnote{Hmm. A quick googling turned up a project out of berkeley combining
OpenMP and TBB, but performance didn't seem too bad. I'll take another look
tomorrow.} In addition, many of these programming models are implemented as
runtime libraries with only primitive compiler support. As a result, most are
not able to take advantage of the compiler's analysis and optimization
capabilities.  While there has been some recent work on specializing
compilers to reason about parallel programming libraries~\cite{Moss_2016},
fragmentation among models presents a significant challenge for compiler
writers, which must choose a single model, or multiple intermediate forms, to
use for analysis and optimization. An ideal solution would be a form of
common representation of parallel semantics and constructs.

@@

Arguably the most ubiquitous parallel programming model for on-node parallelism
in HPC, OpenMP has grown from being a way to implement parallel loops into a
large and complex specification for parallelism. This paper focuses on a 
relatively recent addition to the OpenMP specification: the \texttt{task}
and \texttt{taskwait} constructs. A detailed description of these constructs
is presented in Section~\ref{Sec:OpenMP} and a more complete description can 
be found in the OpenMP specification~\cite{openmp}.

@@

In this section we further discuss a significant challenge in HPC:
fragmentation of parallel programming models. There are many ways to write
parallel programs~\cite{openmp, qthreads, chapel, cilk, kokkos, legion, upc, mpi}
\pmnote{I would add MPI and UPC/PGAS here too}. From libraries, to language
extensions, to standalone languages, to combinations of the above,
fragmentation of parallel programming models is a problem that is only getting
worse. This means that when writing tooling, optimizations, or analyses, one
has to choose which model to target as well as understand points of contention
and interoperability between them. A common difficulty among all of these is
that reasoning about parallel programs is hard and compilers can potentially
help reduce this burden. This is exactly analogous the problem LLVM solves:
compiler optimizations are hard, so sharing them is valuable. Tapir is a
demonstration that the the LLVM approach to compilation of serial code can be
effectively extended to help reason about parallel code.

@@

Initially, the cross-vendor OpenMP~\cite{openmp} shared memory programming model 
focused on the execution of data parallelism by a cooperating team of threads, 
e.g., dividing the iterations of a loop among the threads. Version 3.0 of the 
OpenMP API specification introduced support for lightweight asynchronous tasks, 
designated by the application developer and scheduled onto the team of threads 
by the OpenMP run time implementation.  The \texttt{task} construct applied to 
a structured block of code creates an explicit task, and the \texttt{taskwait} 
construct waits for completion of all tasks generated by the current task.

@@

A seperate but related area in LLVM extensions is the area of using memory
